---
title: "synthesis"
author: "SPT"
date: "2021_10_21"
output: github_document
---

### Step 1: Load packages and data

New and improved. Use `add.rows` function to add multiple effect sizes to the same study. The database is organized chronologically and new rows are added to the end of the database, such that row id's retain their meaning throughout. If you find this, I hope you enjoy it.

```{r, echo = `FALSE`}
# Install packages
#devtools::install_github("MathiasHarrer/dmetar")
#install.packages("esc")

# Load packages
library(tidyverse)
library(meta)
library(metafor)
library(esc)
#library(dmetar)

 # Reset global environment
rm(list = ls())

# Change working directory
setwd("/Users/simontye/Research/Projects/Meta_Analysis/2021_MetaAnalysis")

# Load Web of Science data (WoS)
wos.1 <- read.csv(file = "data/WoS/savedrecs.csv", head = TRUE, sep = ",")
wos.2 <- read.csv(file = "data/WoS/savedrecs-2.csv", head = TRUE, sep = ",")
wos.3 <- read.csv(file = "data/WoS/savedrecs-3.csv", head = TRUE, sep = ",")
wos.4 <- read.csv(file = "data/WoS/savedrecs-4.csv", head = TRUE, sep = ",")
wos.5 <- read.csv(file = "data/WoS/savedrecs-5.csv", head = TRUE, sep = ",")

# Function to insert rows for papers with multiple effect sizes
add.rows <- function(data, newrow, r) {
  data[seq(r + 1, nrow(data) + 1),] <- data[seq(r, nrow(data)),]
  data[r, ] <- newrow
  data
}
```

### Step 2: Format data

```{r echo = `FALSE`}
# Combine WoS data
wos <- wos.1 %>%
  rbind(wos.2, wos.3, wos.4, wos.5) %>%
  rename_all(tolower) %>%
  dplyr::select(., c(publication.year, article.title, authors,
                     source.title, book.series.title, abstract)) %>%
  dplyr::rename(title   = article.title,
                journal = source.title,
                book    = book.series.title,
                year    = publication.year) %>%
  arrange(year, title) %>%
  mutate_all(na_if,"") %>%
  mutate(include       = as.character(NA),
         comment       = as.character(NA),
         focal_species = as.character(NA),
         focal_n       = as.character(NA),
         comp_species  = as.character(NA),
         comp_n        = as.character(NA),
         pred_species  = as.character(NA),
         pred_n        = as.character(NA),
         interaction   = as.character(NA),
         type          = as.character(NA),
         location      = as.character(NA),
         lat           = as.character(NA),
         long          = as.character(NA),
         n             = as.character(NA),
         mean          = as.character(NA),
         sd            = as.character(NA),
         id            = row_number(),
         es.id         = as.character(NA),
         es            = as.character(NA),
         es.se         = as.character(NA),
         es.v          = as.character(NA),
         es.ci.upper   = as.character(NA),
         es.ci.lower   = as.character(NA),
         es.weight     = as.character(NA)) %>%
  dplyr::select(., c(id, year, title, authors,
                     journal, book, abstract,
                     include, comment,
                     focal_species, focal_n,
                     comp_species, comp_n,
                     pred_species, pred_n,
                     type, interaction,
                     location, lat, long,
                     n, mean, sd, es.id,
                     es, es.se, es.v,
                     es.ci.upper, es.ci.lower,
                     es.weight))

# Remove objects
#rm(wos.1, wos.2, wos.3,
#   wos.4, wos.5)
```

### Step 2: Enter data for each article and calculate effect size

```{r, 1, echo = `FALSE`}
### Article 1
wos[1, 8:9] <- c(
  # include
  "N",
  # comment
  "theory")
```

```{r, 2, echo = `FALSE`}
### Article 2
wos[2, 8:9] <- c(
  # include
  "N",
  # comment
  "theory")
```

```{r, 3, echo = `FALSE`}
### Article 3

# Rana sylvatica, no competition
wos[3, 8:24] <- c(# include, comment
                  "Y", "field",
                  # focal_species, n
                  "Rana sylvatica", 150,
                  # comp_species, n
                  "None", 0,
                  # pred_species, n
                  "None", 0,
                  # type, interaction
                  "no_comp", "survivorship",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA, NA,
                  # n, mean, sd
                  2, mean(c(8.6, 22.1)), sd(c(8.6, 22.1)),
                  # effect size id
                  "3_1")

# Rana sylvatica, control
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include, comment
                  "Y", "field",
                  # focal_species, n,
                  "Rana sylvatica", 150,
                  # comp_species, n
                  "Rana pipiens", 150,
                  # pred_species, n
                  "None", 0,
                  # type, interaction
                  "comp", "survivorship",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat, long
                  NA, NA,
                  # n, mean, sd
                  3, mean(c(1.4, 27.6, 4.3)), sd(c(1.4, 27.6, 4.3)),
                  # effect size id
                  "3_1"),
                # Add to end
                sum(nrow(wos)) + 1) %>%
  filter(!is.na(id))

# Save rows
first  <- 3
second <- nrow(wos)

# Calculate effect size (d)
es.3.1 <- esc_mean_sd(grp1m  = as.numeric(wos[c(second),  c("mean")]),
                      grp1sd = as.numeric(wos[c(second),  c("sd")]),
                      grp1n  = as.numeric(wos[c(second),  c("n")]),
                      grp2m  = as.numeric(wos[c(first), c("mean")]),
                      grp2sd = as.numeric(wos[c(first), c("sd")]),
                      grp2n  = as.numeric(wos[c(first), c("n")]),
                      study = "3_1")

# Add effect size to main df
wos[c(first, second), 25:30] <- c(es.3.1$es, es.3.1$se,
                                  es.3.1$var, es.3.1$ci.lo,
                                  es.3.1$ci.hi, es.3.1$w)
```

### Step 3: Notes

```{r, echo = `FALSE`}

# Rana sylvatica, control
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana sylvatica",
                  # focal_n
                  300,
                  # comp_species
                  "None",
                  # comp_n
                  0,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "control",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  3,
                  # mean,
                  mean(c(4.0, 29.6, 1.0)),
                  # sd
                  sd(c(4.0, 29.6, 1.0))),
                # Add to end
                sum(nrow(wos) + 1))

# Rana sylvatica, control
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana sylvatica",
                  # focal_n
                  600,
                  # comp_species
                  "None",
                  # comp_n
                  0,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "control",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  3,
                  # mean,
                  mean(c(16.2, 21.0, 5.2)),
                  # sd
                  sd(c(16.2, 21.0, 5.2))),
                # Add to end
                sum(nrow(wos) + 1))

# Rana sylvatica, control
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana sylvatica",
                  # focal_n
                  1200,
                  # comp_species
                  "None",
                  # comp_n
                  0,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "control",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  3,
                  # mean,
                  mean(c(17.7, 15.3, 0.5)),
                  # sd
                  sd(c(17.7, 15.3, 0.5))),
                # Add to end
                sum(nrow(wos) + 1))

################################################
# Rana sylvatica, competition
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana sylvatica",
                  # focal_n
                  150,
                  # comp_species
                  "Rana pipiens",
                  # comp_n
                  150,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "competition",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  3,
                  # mean,
                  mean(c(1.4, 27.6, 4.3)),
                  # sd
                  sd(c(1.4, 27.6, 4.3))),
                # Add to end
                sum(nrow(wos) + 1))

# Rana sylvatica, competition
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana sylvatica",
                  # focal_n
                  300,
                  # comp_species
                  "Rana pipiens",
                  # comp_n
                  300,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "competition",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  2,
                  # mean,
                  mean(c(7.3, 25.6)),
                  # sd
                  sd(c(7.3, 25.6))),
                # Add to end
                sum(nrow(wos) + 1))

# Rana sylvatica, competition
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana sylvatica",
                  # focal_n
                  600,
                  # comp_species
                  "Rana pipiens",
                  # comp_n
                  600,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "competition",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  2,
                  # mean,
                  mean(c(12.6, 17.0)),
                  # sd
                  sd(c(12.6, 17.0))),
                # Add to end
                sum(nrow(wos) + 1))

# Rana sylvatica, competition
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana sylvatica",
                  # focal_n
                  1200,
                  # comp_species
                  "Rana pipiens",
                  # comp_n
                  1200,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "competition",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  2,
                  # mean,
                  mean(c(9.1, 21.6)),
                  # sd
                  sd(c(9.1, 21.6))),
                # Add to end
                sum(nrow(wos) + 1))

################################################
# Rana pipiens, control
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana pipiens",
                  # focal_n
                  150,
                  # comp_species
                  "None",
                  # comp_n
                  0,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "control",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  3,
                  # mean,
                  mean(c(6.8, 6.4, 2.8)),
                  # sd
                  sd(c(6.8, 6.4, 2.8))),
                # Add to end
                sum(nrow(wos) + 1))

# Rana pipiens, control
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana pipiens",
                  # focal_n
                  300,
                  # comp_species
                  "None",
                  # comp_n
                  0,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "control",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  3,
                  # mean,
                  mean(c(4.7, 32.6, 0.3)),
                  # sd
                  sd(c(4.7, 32.6, 0.3))),
                # Add to end
                sum(nrow(wos) + 1))

# Rana pipiens, control
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana pipiens",
                  # focal_n
                  600,
                  # comp_species
                  "None",
                  # comp_n
                  0,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "control",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  3,
                  # mean,
                  mean(c(3.7, 23.3, 1.2)),
                  # sd
                  sd(c(3.7, 23.3, 1.2))),
                # Add to end
                sum(nrow(wos) + 1))

# Rana pipiens, control
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana pipiens",
                  # focal_n
                  1200,
                  # comp_species
                  "None",
                  # comp_n
                  0,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "control",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  3,
                  # mean,
                  mean(c(8.8, 11.6, 0.4)),
                  # sd
                  sd(c(8.8, 11.6, 0.4))),
                # Add to end
                sum(nrow(wos) + 1))

################################################
# Rana pipiens, competition
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana pipiens",
                  # focal_n
                  150,
                  # comp_species
                  "Rana sylvatica",
                  # comp_n
                  150,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "competition",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  2,
                  # mean,
                  mean(c(9.5, 30.5)),
                  # sd
                  sd(c(9.5, 30.5))),
                # Add to end
                sum(nrow(wos) + 1))

# Rana pipiens, competition
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana pipiens",
                  # focal_n
                  300,
                  # comp_species
                  "Rana sylvatica",
                  # comp_n
                  300,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "competition",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  2,
                  # mean,
                  mean(c(9.2, 23.6)),
                  # sd
                  sd(c(9.2, 23.6))),
                # Add to end
                sum(nrow(wos) + 1))

# Rana pipiens, competition
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana pipiens",
                  # focal_n
                  600,
                  # comp_species
                  "Rana sylvatica",
                  # comp_n
                  600,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "competition",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  2,
                  # mean,
                  mean(c(10.3, 15.0)),
                  # sd
                  sd(c(10.3, 15.0))),
                # Add to end
                sum(nrow(wos) + 1))

# Rana pipiens, competition
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field",
                  # focal_species
                  "Rana pipiens",
                  # focal_n
                  1200,
                  # comp_species
                  "Rana sylvatica",
                  # comp_n
                  1200,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "competition",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  2,
                  # mean,
                  mean(c(4.8, 7.1)),
                  # sd
                  sd(c(4.8, 7.1))),
                # Add to end
                sum(nrow(wos) + 1))

################################################
# Rana sylvatica, control
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field, open",
                  # focal_species
                  "Rana sylvatica",
                  # focal_n
                  "High density",
                  # comp_species
                  "None",
                  # comp_n
                  0,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "control",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  2,
                  # mean,
                  mean(c(25.7, 32.4)),
                  # sd
                  sd(c(25.7, 32.4))),
                # Add to end
                sum(nrow(wos) + 1))

# Rana sylvatica, control
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field, open",
                  # focal_species
                  "Rana sylvatica",
                  # focal_n
                  "Low density",
                  # comp_species
                  "None",
                  # comp_n
                  0,
                  # pred_species
                  "None",
                  # pred_n
                  0,
                  # interaction
                  "survivorship",
                  # type
                  "control",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  2,
                  # mean,
                  mean(c(29.0, 35.2)),
                  # sd
                  sd(c(29.0, 35.2))),
                # Add to end
                sum(nrow(wos) + 1))

################################################
# Rana sylvatica, competition+predation
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field, open",
                  # focal_species
                  "Rana sylvatica",
                  # focal_n
                  "High density",
                  # comp_species
                  "Rana pipiens",
                  # comp_n
                  "High density",
                  # pred_species
                  "Macroinvertebrates",
                  # pred_n
                  "Unknown",
                  # interaction
                  "survivorship",
                  # type
                  "comp+pred",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  2,
                  # mean,
                  mean(c(1.4, 2.3)),
                  # sd
                  sd(c(1.4, 2.3))),
                # Add to end
                sum(nrow(wos) + 1))

# Rana sylvatica, competition+predation
wos <- add.rows(wos,
                c(wos[3, 1:7],
                  # include
                  "Y",
                  # comment
                  "field, open",
                  # focal_species
                  "Rana sylvatica",
                  # focal_n
                  "Low density",
                  # comp_species
                  "Rana pipiens",
                  # comp_n
                  "Low density",
                  # pred_species
                  "Macroinvertebrates",
                  # pred_n
                  "Unknown",
                  # interaction
                  "survivorship",
                  # type
                  "comp+pred",
                  # location
                  "University of Michigan, ES George Reserve, Pinckney",
                  # lat
                  NA,
                  # long
                  NA,
                  # n
                  2,
                  # mean,
                  mean(c(2.6, 0.6)),
                  # sd
                  sd(c(2.6, 0.6))),
                # Add to end
                sum(nrow(wos) + 1))


```

```{r, echo = `FALSE`}
# Subset current data
current <- wos %>%
  filter(!is.na(include)) %>%
  filter(include == "Y")



# Filter control group
control <- current %>%
  filter(type == "control") %>%
  arrange(focal_species, focal_n)

# Filter competition group
comp <- current %>%
  filter(type == "competition") %>%
  arrange(focal_species, focal_n)

# Filter predation group
pred <- current %>%
  filter(type == "predation") %>%
  arrange(focal_species, focal_n)

# Filter comp/pred group
comp.pred <- current %>%
  filter(type == "comp+pred") %>%
  arrange(focal_species, focal_n)

# Calculate effect sizes
es.l <- esc_mean_sd(grp1m  = as.numeric(comp$mean),
                    grp1sd = as.numeric(comp$sd),
                    grp1n  = as.numeric(comp$n),
                    grp2m  = as.numeric(control$mean),
                    grp2sd = as.numeric(control$sd),
                    grp2n  = as.numeric(control$n))

# Convert to df
es.df <- as.data.frame(effect.sizes)

# Test plot
ggplot() +
  geom_point(es.df, mapping = aes(x = es,
                                  y = study)) +
  geom_errorbar(es.df, mapping = aes(xmin = (es - se),
                                     xmax = (es + se),
                                     y = study,
                                     width = 0.1)) +
  labs(x = "Effect size (d)", y = "Study")

## Calculate values we need for the formulas
#s1 <- sd(x1)
#s2 <- sd(x2)
#n1 <- 20
#
## Calculate s_pooled
#s_pooled <- sqrt(
#  (((n1-1)*s1^2) + ((n2-1)*s2^2))/
#    ((n1-1)+(n2-1))
#)
## Calculate the standard error
#se <- s_pooled*sqrt((1/n1)+(1/n2))
#se

```

